Class Scheduling Problem

### 2024-05-13 homework1/paper.tex
last modification

\section{Introduction}
\subsection{Problem Declaration}

@@

For instance, in instances where an instructor's preferences are violated within 
a particular time slot, the algorithm should endeavor to substitute instructor with 
another whose preferences remain unviolated for that time slot. Moreover, the 
algorithm should endeavor to substitute the time slot with another or substitute
the room with anothers that have the same capacity in total. This approach
was found more difficult to implement. When tring to implement the first two
substitutes, the algorithm was not able to find a solution that satisfies all the
soft constraints and it took quite a lot of time to find a partial solution due to
a large number of constraints that should be checked while creating the neighbors.
On the other hand, this approach will always find a partial solution that 
satisfies all the hard constraints. Thus, this one is recommended in cases 
when the soft constraints are not that important.

@@

As previously mentioned, the Random Restart Hill Climbing algorithm is the 
most suitable for this approach. The algorithm is initialized with an empty
schedule and generates neighbors that adhere to the hard constraints. The
algorithm iteratively explores the neighborhood of the current state, selecting
randomly one of the neighbors that satisfies the most constraints. This process continues until
a local optimum is reached, at which point the algorithm restarts the search.
The algorithm terminates when a solution that
satisfies all constraints is found or when a specified number of iterations / 
restarts have been completed. A pseudocode of the algorithm is presented in Algorithm
\ref{alg1}.

@@

\State $state = random.choice(from\ neighbors\ one\ of\ the\ neighbors\ with$ \\
\hspace*{2em}$\text{minimum\ number\ of\ constraints\ unsatisfied)}$
\EndWhile

@@

The pseudocode of the algorithm is presented in Algorithm \ref{alg2}.

@@

\subsection{Results}
Pentru fiecare dintre algoritmii prezentați mai sus, am folosit \textit{hyperfine}
pentru a evalua timpul de rulare al acestora și am salvat datele în câte un fișier
MD. Excepție a făcut algoritmul Brute Force, căruia i-am calculat timpul de rulare
doar pentru testele 1-2, 3, 9-14, 15-16, deoarece pentru celelalte teste timpul de
rulare era prea mare.

Pentru fiecare set de date, am făcut un grafic cu timpul de rulare al fiecărui
algoritm, iar rezultatele sunt prezentate în Figurile \ref{fig:empty_graphs} - \ref{fig:fixed_edge_number_graphs}.

De asemenea, în tabelul \ref{tab:results} sunt prezentate rezultatele obținute
pentru fiecare set de date în raport cu numărul cromatic teoretic al grafurilor (
    acest număr a putut fi precalculat doar pentru testr 1-26
).
Pentru anumite teste care au fost prea costisitoare în timp pentru algoritmul Brute Force, am
lăsat liber câmul respectiv.
\pagebreak

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./graphics/fig1.png}
\caption{Graficele timpului de rulare pentru grafuri goale}
\label{fig:empty_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./graphics/fig2.png}
\caption{Graficele timpului de rulare pentru grafuri complete}
\label{fig:complete_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./graphics/fig3.png}
\caption{Graficele timpului de rulare pentru grafuri bipartite complete}
\label{fig:bipartite_complete_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./graphics/fig4.png}
\caption{Graficele timpului de rulare pentru arbori binari}
\label{fig:binary_tree_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./graphics/fig5.png}
\caption{Graficele timpului de rulare pentru grafuri planare}
\label{fig:planar_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./graphics/fig6.png}
\caption{Graficele timpului de rulare pentru grafuri cu număr de muchii fixat}
\label{fig:fixed_edge_number_graphs}
\end{figure}

\begin{table}[!]
\centering
\caption{Rezultatele obținute pentru fiecare set de date}
\label{tab:results}
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
\textbf{Nr. set} & \textbf{Nr. noduri} & \textbf{Nr. muchii} & \textbf{Nr. cromatic} & \textbf{Brute Force} & \textbf{Greedy} & \textbf{DSatur} & \textbf{RLF} \\ \hline
\textbf{1} & 887 & 0 & 1 & 1 & 1 & 1 & 1 \\ \hline 
\textbf{2} & 1384 & 0 & 1 & 1 & 1 & 1 & 1 \\ \hline
\textbf{3} & 10 & 45 & 10 & 10 & 10 & 10 & 10 \\ \hline 
\textbf{4} & 50 & 1225 & 50 & 50 & 50 & 50 & 50 \\ \hline 
\textbf{5} & 100 & 4950 & 100 & - & 100 & 100 & 100 \\ \hline 
\end{tabular}
\end{table}

\subsection{Interpretarea rezultatelor}
În raport cu timpul de execuție, avem următoarele observații:
\begin{itemize}
    \item Algoritmul Brute Force este cel mai lent, \textbf{nu} se folosește
    în practică deloc.
    \item Algoritmul Greedy pe toate tipurile de grafuri este cel mai rapid.
    \item Algoritmul DSatur este mai rapid decât algoritmul Brute Force, dar
    mai lent decât algoritmul Greedy.
    \item Algoritmul RLF pe grafuri goale este mai lent și decât algoritmul
    Brute Force, pe grafuri planare și arbori binari este mai lent decât DSatur, 
    iar pe restul de grafuri este între Greedy și DSatur (aici au apărut cazuri
    care în teorie nu sunt adevărate din cauza complexității agoritmului (vezi \ref{tabel_complexitate}).
    Acest lucru se datorează faptului că nu s-au putut genera teste suficient de
    mari și rulate de un număr suficient de ori pentru a se observa diferența dintre algoritmi.)

\end{itemize}

În raport cu numărul cromatic, se poate observa cu ușurință faptul că algoritmul
Brute Force este un algoritm exact, iar restul sunt aproximări (deși în teste 
nu se observă o diferență între numărul cromatic teoretic și rezultatul de la algoritmul DSatur, dacă
am fi avut un număr mai mare de teste cu numărul cromatic precalculat, ar fi apărut diferența).
Dintre toți1 algoritmii euristici, algoritmul DSatur este cel mai exact.

\section{Concluzii}
În cadrul acestei teme am observat 4 algoritmi diferiți de colorare a grafurilor.

Se poate afirma că nu există o soluție complet generală pentru colorarea grafurilor.
Fiecare din cele prezentate mai sus vin cu avantajele și dezavantajele lor.

Algoritmul Brute Force este cel mai lent, dar este exact, algoritmul Greedy este
cel mai rapid, dar nu este exact, algoritmul DSatur deși este mai lent decât Greedy,
este cel mai exact dintre algoritmi, iar algoritmul RLF este mai lent decât Brute Force și 
DSatur, dar este aproximativ exact.

În practică:
\begin{itemize}
    \item Dacă lucrați cu grafice bipartite sau ciclice, algoritmul Greedy nu este
    suficient de eficient; pe de altă parte, algoritmul RLF este destul de exact și total
    recomandat pentru aceste cazuri.
    \item Dacă prioritar este timpul, se poate înclina spre utilizarea algoritmul Greedy.
    \item Dacă prioritatea este un număr mai apropiat de cel cromatic, trebuie
    să se utilizeze algoritmul DSatur.

\end{itemize}
\pagebreak
% ------------------------------------------------------------------------------

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}

    \bibitem{1}
    J. Bondy and U. Murty, Graph Theory - Graduate Texts in Mathematic.
    Springer, 2008.
    \bibitem{2}
    Z. \`{A}d\`{a}m Mann and A. Szajk\`{o}, “Average-case complexity of backtrack
    search for coloring sparse random graphs,” Journal of Computer and
    System Sciences, vol. 79, no. 8, pp. 1287–1301, 2013.
    \bibitem{3}
    “Backtrack: An o(1) expected time algorithm for the graph coloring
    problem,” Information Processing Letters, vol. 18, no. 3, pp. 119–121,
    1984.
    \bibitem{4}
    N. Barnier and P. Brisset, “Graph coloring for air traffic flow management,” 
    Annals of Operations Research, vol. 130, 03 2002.
    \bibitem{5}
    “The application of a graph coloring method to an examination scheduling problem,” 
    Institute for Operations Research and the Management
    Sciences (INFORMS), vol. 11, no. 5.
    \bibitem{6}
    \href{http://www.ijimt.org/papers/359-D0129.pdf}{University Exam Scheduling System Using Graph
    Coloring Algorithm and RFID Technology}
    \bibitem{7}
    A. Murat and B. Nurdan, “A performance comparison of graph coloring algorithms,” 
    International Conference on Advanced Technology
    Sciences (ICAT’16), vol. 4, pp. 1–19, 12 2016
    \bibitem{8}
    Z. Mann, “Complexity of coloring random graphs: An experimental
    study of the hardest region,” Journal of Experimental Algorithmics,
    vol. 23, pp. 1–19, 03 2018
    \bibitem{9}
    \href{https://acs-aa-challenge.github.io/acs-aa-challenge/18-np-2-colouring/}{Enunțul problemei}
    \bibitem{10}
    M. Garey and D. Johnson, Computer and Intractability: A Guide to the
    Theory of NP-Completeness, 01 1979.
    \bibitem{11}
    D. Porumbel, J.-K. Hao, and P. Kuntz, “An evolutionary approach with
    diversity guarantee and well-informed grouping recombination for graph
    coloring,” Computers Operations Research, vol. 37, pp. 1822–1832, 10
    2010.
    \bibitem{12}
    L. Ouerfelli and H. Bouziri, “Greedy algorithms for dynamic graph coloring,” 2011 International Conference on Communications, Computing
    and Control Applications, CCCA 2011, 03 2011.
    \bibitem{13}
    \`{A}. E. Eiben, J. K. Van Der Hauw, and J. I. van Hemert, “Graph coloring
    with adaptive evolutionary algorithms,” Journal of Heuristics, vol. 4,
    no. 1, pp. 25–46, 1998.
    \bibitem{14}
    D. Br\`{e}laz, “New methods to color the vertices of a graph,”  Commun.
    ACM, vol. 22, pp. 251–256, 04 1979.
    \bibitem{15}
    R. Janczewski, K. Manuszewski, and K. Piwakowski, “The smallest
    hard-to-color graph for algorithm dsatur,” Discrete Mathematics, vol.
    236, pp. 151–165, 06 2001.
    \bibitem{16}
    \href{https://github.com/ifsmirnov/jngen}{GitHub - Jngen: preparing tests made simpler}
    \bibitem{}
    \href{https://en.wikipedia.org/wiki/Graph_coloring}{Graph coloring - Wikipedia}
    \bibitem{}
    \href{https://citeseerx.ist.psu.edu/document?repid=rep1&type=pdf&doi=203a7b17267a28a06808bfb3b0b9571e32d15503}{A Comparison of Parallel Graph Coloring Algorithms}
    \bibitem{}
    \href{https://dergipark.org.tr/en/download/article-file/254140}{A Performance Comparison of Graph Coloring Algorithms}
    \bibitem{}
    \href{https://curs.upb.ro/2022/mod/folder/view.php?id=77105}{Moodle - Analiza algoritmilor}
    % 
    \end{thebibliography}
\end{document}

>>>>

% ------------------------------------------------------------------------------
\pagebreak
\section{Introduction}
\subsection{Problem Declaration}

@@

For instance, in instances where an instructor\'s preferences are violated within 
a particular time slot, the algorithm should endeavor to substitute instructor with 
another whose preferences remain unviolated for that time slot. Moreover, the 
algorithm should endeavor to substitute the time slot with another or substitute
the room with anothers that have the same capacity in total. This approach
was found more difficult to implement. When tring to implement the first two
substitutes, the algorithm was not able to find a solution that satisfies all the
soft constraints and it took quite a lot of time to find a partial solution due to
a large number of constraints that should be checked while creating the neighbors.
On the other hand, this approach will always find a partial solution that 
satisfies all the hard constraints. Thus, this one is recommended in cases 
when the soft constraints are not that important.

@@

As previously mentioned, the Random Restart Hill Climbing algorithm is the 
most suitable for this approach. The algorithm is initialized with an empty
schedule and generates neighbors that adhere to the hard constraints. The
algorithm iteratively explores the neighborhood of the current state, selecting
randomly one of the neighbors that satisfies the most constraints. This process continues until
a local optimum is reached, at which point the algorithm restarts the search.
The algorithm terminates when a solution that
satisfies all constraints is found or when a specified number of iterations / 
restarts have been completed. A pseudocode of the algorithm is presented in Algorithm 1.

@@

\State $state = random.choice(from\ neighbors\ one\ of\ the\ neighbors\ with$ \\
\hspace*{2em}$minimum\ number\ of\ constraints\ unsatisfied)$
\EndWhile

@@

The pseudocode of the algorithm is presented in Algorithm 2.

@@

\subsection{Results}
For the Hill-Climbing algorithm, the number of restarts was established at 100,
while the maximum number of iterations for all tests was capped at 1000. In the 
case of the A* algorithm, a weight of 100 was applied.

Each test involved the execution of both algorithms. The ensuing tables depict 
the outcomes across various categories: the count of unsatisfied soft 
constraints, the time taken to attain a solution, the total number of states 
explored during the search, and the iterations needed to reach a solution. 
Additionally, for the Hill Climbing algorithm, the number of restarts executed 
until a solution was obtained is delineated.

The outcomes for the Hill Climbing algorithm are categorized based on the 
number of restarts.

Moreover, the individual results for each test are presented in the following
tables 1 and 2.

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./images/iterations.png}
\caption{Number of iterations required to find the solution}
\label{fig:empty_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./images/states.png}
\caption{Total number of states explored during the search}
\label{fig:complete_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./images/times.png}
\caption{Time required to find the solution}
\label{fig:bipartite_complete_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./images/constraints.png}
\caption{Number of soft constraints unsatisfied}
\label{fig:binary_tree_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./images/restarts.png}
\caption{Number of restarts until a solution was found for Hill Climbing}
\label{fig:planar_graphs}
\end{figure}

\begin{table}[!]
\centering
\caption{Results obtained for the Hill Climbing algorithm}
\label{tab:results_hc}
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
\textbf{No. set} & \textbf{No. iterations} & \textbf{No. states} & \textbf{Time mm.ss.ms} & \textbf{No. unsatisfied constraints} & \textbf{No. restars} \\ \hline
\textbf{1} & 106 & 9989 & 0.4.317 & 0 & 3  \\ \hline 
\textbf{2} & 71 & 53875 & 0.34.053 & 0 & 1 \\ \hline
\textbf{3} & 76 & 72621 & 0.77.753 & 0 & 1  \\ \hline 
\textbf{4} & 5889 & 1228027 & 10.40.491 & 6 & 100 \\ \hline 
\textbf{5} & 2221 & 1603927 & 22.47.289 & 5 & 20 \\ \hline 
\end{tabular}
\end{table}

\begin{table}[!]
\centering
\caption{Results obtained for the A Star algorithm}
\label{tab:results_astar}
\begin{tabular}{|c|c|c|c|c|}
\hline
\textbf{No. set} & \textbf{No. iterations} & \textbf{No. states} & \textbf{Time mm.ss.ms} & \textbf{No. unsatisfied constraints} \\ \hline
\textbf{1} & 146 & 17073 & 0.8.537 & 0  \\ \hline
\textbf{2} & 158 & 183159 & 2.12.937 & 0  \\ \hline
\textbf{3} & 2883 & 2171276 & 52.49.856 & 0  \\ \hline
\textbf{4} & 91 & 12368 & 0.8.172 & 0  \\ \hline
\textbf{5} & 2542 & 1385784 & 27.15.878 & 0  \\ \hline
\end{tabular}
\end{table}

\subsection{Observations}

The Hill Climbing algorithm exhibits faster performance compared to the A* 
algorithm primarily due to its local search nature. The Hill Climbing algorithm
is more efficient in terms of the number of iterations required to reach a
solution.

The A* algorithm, while slower, is more effective in terms of the number of
unsatisfied constraints. The A* algorithm consistently yields solutions that
satisfy all constraints, whereas the Hill Climbing algorithm occasionally
encounters unsatisfied constraints.

Recognizing the prolonged duration required to attain a solution for 
orar\_bonus\_exact using the Hill Climbing algorithm, the number of 
restarts was reduced to 20.

The A Star algorithm is more suitable for scenarios similar to orar\_mic\_exact,
orar\_mediu\_relaxat, and orar\_constrans\_incalcat, where the number of
neighbors generated is smaller. In contrast, the Hill Climbing algorithm is
more appropriate for scenarios similar to orar\_mare\_relaxat and orar\_bonus\_exact,
where the number of neighbors generated is larger.

\section{Conclusions}

In summary, within our implementation of the Class Scheduling Problem, 
the Hill Climbing algorithm outperforms the A* algorithm in terms of efficiency.
A* achieves slower solution discovery, but with no unsatisfied constraints. 
Conversely, Hill Climbing is faster, but may occasionally encounter unsatisfied
constraints. The Hill Climbing algorithm is more suitable for scenarios where
the soft constraints are less significant and the time required to find a
solution is a critical factor. In contrast, the A* algorithm is more appropriate
for scenarios where the soft constraints are more significant and the quality of
the solution is paramount.

\pagebreak
% ------------------------------------------------------------------------------

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
    \bibitem{}
    \href{https://en.wikipedia.org/wiki/A*_search_algorithm#Complexity}{A* search algorithm - Wikipedia}
    \bibitem{}
    \href{https://en.wikipedia.org/wiki/Hill_climbing}{Hill Climbing - Wikipedia}
    \bibitem{}
    \href{https://curs.upb.ro/2023/course/view.php?id=13749}{Moodle - Artifical Intelligence Course}
    \bibitem{}
    \href{https://www.researchgate.net/profile/Kian-Pokorny/publication/262277120_Multiple_constraint_satisfaction_problems_using_the_A-star_A_search_algorithm_classroom_scheduling_with_preferences/links/5e57f530299bf1bdb8408b03/Multiple-constraint-satisfaction-problems-using-the-A-star-A-search-algorithm-classroom-scheduling-with-preferences.pdf}{Multiple constraint satisfaction problems using the A-star search algorithm}
    \bibitem{}
    \href{https://d1wqtxts1xzle7.cloudfront.net/68245993/14443-libre.pdf?1626958962=&response-content-disposition=inline%3B+filename%3DCourses_timetabling_based_on_hill_climbi.pdf&Expires=1714507660&Signature=JDd-gaDdE4E7s5P66Op535FyEuysvPoA5o4fcwak4NWT1sQNtVh50ooYa9eSrHu~aHR-Ulabenz910uDjozKpgHivCFkhipTT5bJJL9MensZTSib8TBVzGtBU9lycqCVwilt21Udi0DMgfhERDSgCcR4vO8hmlFlxOLg1exD1fNLE8kxt4S0H2Qfb3nQEHu3kLsicuJNk8wfF~eWpEDUlWsjW9J9IpxuLWP0OICQ7ZReWtx37dHnPN2PTVmFs815NMNi-jyLV0gWWBTAOL7Hn7A5VSg-l~s2e8RAoCjPAYQDRnnznIzOElRy2nTsgkW3mZDzlsntJ3lFur2KquwqNg__&Key-Pair-Id=APKAJLOHF5GGSLRBV4ZA}{Courses timetabling based on hill climbing algorithm}

    \end{thebibliography}
\end{document}
